---
title: "Regresion Models Course Project"
author: "Hernan Solano"
date: "19/5/2018"
output: 
  pdf_document: 
    fig_caption: yes
    fig_height: 4
    fig_width: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(GGally)
library(ggthemes)
```

# Executive Summary
This is the final project for the Regression Models course of the Data Science Specialization. We are going to use the mtcars dataset from the default data of the R software in order to determing the relationship between the nature of the miles per galon (MPG) and the transmission or am variable (0 = automatic, 1 = manual) of a series of vehicles. We perform some Exploratory Data Analisis and then chose an appropiate regression model in order to quantify the difference of MPG of each transmission type.

#Exploratory Data Analysis
    
    First, lets identify the variable types and how the table looks like:

```{r cars}
head(mtcars,3)
```
    Since our focus is on the mpg and am variables the following code shows their key measures:

```{r summary1}
round(summary(mtcars$mpg),2)
round(summary(mtcars$am),2)
```
In this case, the range of the mpg variable is (10.4 - 33.9) and its standard deviation is: `r round(sd(mtcars$mpg),2)`. This is consistent with the figure 1 that shows the histogram of the mpg variable. On the other, the am variable only can take as values a 0 or a 1 and 41% of the cars have a manual transmission. 

Furthermore, if we see figure 2, we can see that cars with a manual transmission might have a higher mean mpg value than the vehicules with an automatic transmision.

#Regression Model
We are going to use a linear regression model in order to identify the relationship between these two variables. Then, our outcome will be the mpg variable and the predictor will be the transmission variable as a factor. We add -1 to the model to avoid confussions with the coefficients.
    
```{r model}
    fit <- lm(mpg ~as.factor(am)-1,data = mtcars)
    summary(fit)
    confint(fit)
   
   
```
According to the summary, none of the estimated coefficients is equal to 0 and the confidence intervals confirm it. We ploted the line on figure 3 so we can appreciate the increasing trend between automatic and manual transmisions. Additionally, if we look at the diagnosis plots in figure 4 there are 3 cars with high leverage in comparison with the others, these are: the Toyota Corolla, the Maserati Bora and the Ford Pantera L.

We can also calculate the difference between the coefficients, that is:

Manual transmition coefficient - automatic transmition coefficient

`r round(fit$coeff[2],2)` - `r round(fit$coeff[1],2)` = `r round(fit$coeff[2] - fit$coeff[1],2)`

We can also confirm this difference by running a t-test like this:

```{r test}
    test <- t.test(mpg ~as.factor(am),data = mtcars)
    test$conf.int
    test$estimate
```
This means that with a p-value of `r test$p.value` we can reject the null hypothesis and say that the mpg values of the 2 compared groups are different. We can conclude by saying that cars with manual transmissions have `r round(fit$coeff[2] - fit$coeff[1],2)` mpg more than cars with automatic transmissions.

# Appendix

##Figure 1

This is the histogram of the mpg variable:
```{r figure 1, fig.height=3, fig.width=4,fig.align='center'}
    ggplot(data = mtcars,aes(x = mpg)) + geom_histogram(fill = "#014d64") + 
    theme_economist() + labs(title = "Figure 1: Histogram of mpg")
```

##Figure 2

This is a generalized Pairs Plot using ggpairs:
```{r figure 2,fig.align='center'}
    ggpairs(data = mtcars,columns = c("wt","am","mpg")) + 
    labs( title = "Figure 2: Pairs Plot") +theme_economist()
```

##Figure 3

This is a plot of the regression line modeled:
```{r figure 3, fig.height=3, fig.width=4,fig.align='center'}
    ggplot(data = mtcars,aes(x = am,y = mpg)) + geom_point()+geom_smooth(method = lm) + 
    labs( title = "Figure 3: Regression line") + theme_economist()
```

##Figure 4

Diagnosis plots:
```{r figure 4,fig.height=5,fig.width=6,fig.align='center'}
    par(mfrow = c(2,2))
    plot(fit)
```