---
title: "Peer assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Processing

This is the final assignment in the Reproducible Research course of the Johns Hopkins University's Data Science specilization on Coursera. We are using the storm dataset from the National weather service and you can find the dataset documentation here: https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf

First I have changed the working directory and loaded the tidyverse packages so I can easilly manipulate and plot the data.

```{r setting up}
setwd("~/Github/DataScienceSpecialization/Reproducible Research/Peer assignment")
library(tidyverse)
```
```{r preprocessing, cache=TRUE}
    stormdata <- read_csv("repdata-data-StormData.csv")
```



## Results

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
