---
title: "Peer assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This report aims to identify the most harmful events across all the US, we are going to see which type of events caused more damaged to population's health and properties. We are using the storm dataset from the National weather service and you can find the dataset documentation here: https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf

## Data Processing

First I have changed the working directory and loaded the tidyverse packages so I can easilly manipulate and plot the data.
```{r setting up}
setwd("~/Github/DataScienceSpecialization/Reproducible Research/Peer assignment")
library(tidyverse)
```

Then I loaded the data using the read_csv from the readr package and I cached this operations in order to better work later.

```{r preprocessing, cache=TRUE}
    stormdata <- read_csv("repdata-data-StormData.csv")
```
In orther to successfully identify the most harmful events in terms of population health I grouped the stormdata by the type of event variable (EVTYPE) then for each type of event calulated the total fatalities and injuries caused. 

Following an invented standart we are going to say that an event is very harmful if it has caused more than 199 deaths and/or more than 599 injuries. So I have created a categorical variable that will have two levels "not that harmful" and "very harmful".

```{r question 1, include =TRUE, cache=TRUE}
dataset1 <- group_by(stormdata,EVTYPE) %>% summarize(FATALITIES = sum(FATALITIES),INJURIES = sum(INJURIES)) %>% mutate( health = if_else(FATALITIES >=200 | INJURIES >= 600,"Very Harmful","Not that harmful")) 
```

In orther to identify the most harmful events on the US in terms of property damage first we took into account the property and the crop damage to have a more accurate estimate. In order to determine the total damage we had to first determine the dollar units (e.g. B for billions and K for thousands) from the unit variable, once we have determined the units we multiply the damage value by the apropiate unit, finally, we sum the total property and total crop damage for each type of event.

Then if we use the same approach that before we can classify the most expensive type of events in the US using 1 Billion dollar in total damage as a treshhold so that less than that events are classified as "Not that harmful" and the others as "Very Harmfull"

```{r question2,include =TRUE, cache=TRUE}
dataset2 <- select(stormdata,EVTYPE,PROPDMG,PROPDMGEXP,CROPDMG,CROPDMGEXP) %>%
    filter(PROPDMG > 0 | CROPDMGEXP > 0) %>% 
    mutate(PROPDMGMULT = case_when(
    PROPDMGEXP == "K" ~ 10^3,
    PROPDMGEXP == "M" ~ 10^6,
    PROPDMGEXP == "B" ~ 10^9,
    PROPDMGEXP == "h" ~ 10^2,
    PROPDMGEXP == "m" ~ 10^-3,
    PROPDMGEXP == "2" ~ 2,
    PROPDMGEXP == "3" ~ 3,
    PROPDMGEXP == "4" ~ 4,
    PROPDMGEXP == "5" ~ 5,
    PROPDMGEXP == "6" ~ 6,
    PROPDMGEXP == "7" ~ 7,
    TRUE ~ 1
)) %>% 
    mutate(CROPDMGMULT = case_when(
    CROPDMGEXP == "K" ~ 10^3,
    CROPDMGEXP == "M" ~ 10^6,
    CROPDMGEXP == "B" ~ 10^9,
    CROPDMGEXP == "h" ~ 10^2,
    CROPDMGEXP == "m" ~ 10^-3,
    CROPDMGEXP == "2" ~ 2,
    CROPDMGEXP == "k" ~ 10^3,
    TRUE ~ 1
)) %>% 
    mutate(PROPDMG = PROPDMG*PROPDMGMULT, CROPDMG = CROPDMG*CROPDMGMULT) %>% 
    mutate(TOTALDMG = PROPDMG+CROPDMG) %>% 
    select(EVTYPE,PROPDMG,CROPDMG,TOTALDMG) %>%
    arrange(desc(TOTALDMG)) %>%
    mutate( property = if_else(TOTALDMG >=10^9,"Very Harmful","Not that harmful"))
```
## Results

To show your the most harmful type of events for human health, the next plot shows the name of the event as text and its xy coordinates represent the logarithm of the total number of fatalities and injuries caused. In this sense, the most harmful events will be in the upper right corner, in this case a tornado is the most harmful. 

```{r health,include=TRUE}
ggplot(dataset1, aes(log10(FATALITIES),log10(INJURIES), label = EVTYPE,color = health)) + geom_text(check_overlap = TRUE,size = 2,angle = -65) +labs(title="Most harmful events for human health", x="Log10 of Total number of Fatalities", y="Log10 of Total number of Injuries")
```

The following list shows the 10 most harmful events in the United States organized in descending order first by the total number of Fatalities and then by the number of injuries.
```{r list}
filter(dataset1,health == "Very Harmful") %>% arrange(desc(FATALITIES),desc(INJURIES))
```

On the other hand, to apreciate the most harmful type of events in terms of property damage, the next plot shows the type of events classified by our categorical variable. Unlike before the most harmfull are in the upper part of the plot. In this case we used as xy coordinates the logarithms of the total Property damage and the total damage (property + crop) respectively.

```{r propdamage,include=TRUE,cache=TRUE}
ggplot(dataset2,aes(log10(PROPDMG),log10(TOTALDMG),label = EVTYPE,color=property)) + geom_text(check_overlap = TRUE,size = 2) + labs(title="Most harmful events in terms of Property damage",x = "Log10 of Total Property Damage",y = "Log10 of Total Damage")
```

The following list shows the ten most harmful type of events for the properties in the US:

```{r list2,include=TRUE}
filter(dataset2,property == "Very Harmful")
```